Dark matter haloes: a multistream view

### 2018-03-07 Draft/Chapter5/Source_v2/draft.tex
initial commit

\begingroup
\let\clearpage\relax
%\tableofcontents
\endgroup
\newpage
\end{comment}
%-------------------------------------------------------------------------------------------------------------------------------
\chapter{Dark matter haloes: a multistream view}

@@

\begin{figure*} 
\centering\includegraphics[height=7cm]{Chapter5/Source_v2/fig1.pdf} 
\caption{Dynamical collapse of dark matter in one-dimensional universe: Top panels show the $(\mathbf{p}, \mathbf{x})$ phase-space manifold of the dark matter sheet at redshifts $z_1$, $z_2$, $z_3$ and $z = 0$. Dots represent the dark matter particles. The momentum values are chosen at arbitrary scales. Bottom panels show the corresponding multistream field multistream field $n_{str}(\mathbf{x}, z)$ (red) and density field $\rho(\mathbf{x}, z)$ (gray). }
\label{fig:1d}
\end{figure*}

We begin with a simple illustration showing the formation of a few haloes in a one-dimensional simulation. Dark matter clustering in a (1+1)-dimensional phase-space $(\mathbf{p}, \mathbf{x})$ (where $p$ is the momentum and $x$ is the co-moving Eulerian coordinate) at four successive time steps is shown in the top panels of \autoref{fig:1d}. The lower panels show the corresponding multistream field (\citealt{Shandarin2012} and \citealt{Abel2012}) $n_{str}(\mathbf{x}, z)$ (red) and density field $\rho(\mathbf{x}, z)$ (gray). At $z_1$ (left-most panel), velocity is single-valued in Eulerian co-ordinates shown, except at a small three-stream region near $\mathbf{x} = 5\pi/4$. This is the first instance of multistreaming in the region, which was previously had $n_{str} = 1$ throughout. The interface of $n_{str} = 1$ and $n_{str} = 3$ regions is also the location of the first caustic. On the other hand, the density calculated at a high resolution shows variations, even in the mono-streaming regions. The variations are especially more pronounced around  the caustic (near $\mathbf{x} = 5\pi/4$).

@@

\begin{figure} 
\centering\includegraphics[width=8cm]{Chapter5/Source_v2/fig2.pdf} 
\caption{Multistream field contours: The multistream field is calculated at 8 times the native resolution. Each inner convex blobs (red) surround local multistream maxima inside. Surrounding outer shell(blue) is not convex throughout the surface, and the outermost gray multistream surface displays a filamental geometry.}
\label{fig:full}
\end{figure}

@@

The emphasis of this paper is to demonstrate the use of multistream field in identifying potential dark matter haloes, and not a full statistical analysis of halo properties. For this purpose, we have run simulations at two different mass resolutions (number of particles $N_p = 128^3$ and $256^3$, and respective mass of particles, $m_p = 3.65 \times 10^{10} h^{-1} M_{\odot}$ and $ 4.57 \times 10^{9} h^{-1} M_{\odot}$), with the same periodic side length $L = 100 h^{-1} Mpc$. The gravitational softening length $\epsilon = 20 h^{-1} kpc$ and $10 h^{-1} kpc$ for low and high resolution simulation respectively. The initial conditions are generated by {MUSIC} \citep{Hahn2011a} with the transfer function from \cite{Eisenstein1998a} at a redshift of $z_{ini}= 80$. The $\Lambda$CDM cosmological simulation is run using {GADGET-2} (\citealt{Springel2005} and \citealt{Springel2001}) is similar to the ones used in \cite{Ramachandra2017}. The cosmological parameters used in the simulation are $\Omega_{m}= 0.276$, $\Omega_{\Lambda}= 0.724$, the Hubble parameter, $h = 0.703$, the power spectrum normalization, $\sigma_8 = 0.811$ and the spectral index $n_s= 0.961$.

@@

We intend to identify haloes in the $n_{str}(\mathbf{x})$ field instead of using just the position coordinate data. While the eigenvalue analysis itself is done at a chosen time, the multistream field inherently has data from six-dimensional Lagrangian space $(\mathbf{q}, \mathbf{x})$ that contains the full dynamical information,similar to the phase-space sheet albeit in a different form. Dynamical history that is embedded in the multistream field is crucial in understanding the strength of gravitational binding of the particles. 
A physically motivated distinction between void and gravitationally collapsed regions -- voids are the regions with a single stream -- is a unique feature of multistream analysis (\citealt{Shandarin2012} and \citealt{Ramachandra2017}). Thus the haloes detected from local maxima of the $n_{str}$ field can be ensured to be away from the mono-streaming voids. Methods based on linking-length or density fields may not be able to ensure that  all the particles in haloes are away from voids (as shown for FOF haloes in \citealt{Ramachandra2017}) .

@@

Hessian matrix $\mathbf{H}(f)$ of a scalar field $f$ involves local second-order variations in three orthogonal directions. Each element of the Hessian matrix $\mathbf{H}_{ij}(f)$ (where $i$ and $j$ can be any of $x$, $y$ or $z$ directions) is given by \autoref{eq:Hess}.

\begin{equation}
\label{eq:Hess}
 \mathbf{H}_{ij}(f) = \frac{\partial^2 f}{\partial x_i \partial x_j}
\end{equation}

@@

(i) Every element of Hessian matrices $\mathbf{H}(-n_{str})$, and consequently the eigenvalues $\lambda_i$'s are zero in single-stream voids. Even if the multistream field is a smoothed, the eigenvalues peak at zero. This property is unique to multistream fields. Eigenvalues of Hessians of density \cite{Aragon-Calvo2007}, velocity shear tensor \cite{Libeskind2013} do not peak at zero, and the eigenvalues of deformation tensor are negative in voids as a result of continuity equation (shown in Zel’dovich formalism as well).

@@

{Our goal is to isolate the locations of convex geometries in the multistream flow field. Prospective regions of the halo locations in the web structure are selected by positive definite condition of the Hessian $\mathbf{H}(-n_{str})$: $\lambda_1 > 0$, $\lambda_2 > 0$ and  $\lambda_3 > 0$, or simply the smallest of the eigenvalues, $\lambda_3 > 0$. We also filter out the regions if the multistream values inside do not suggest gravitational collapse into haloes. The sequence of our halo detection framework is listed below:

@@

\item Second order variations of the smoothed $-n_{str}(\mathbf{x})$ is computed at each point in the field. This gives symmetric Hessian matrices for this field whose eigenvalues are real. Ordered eigenvalues of the Hessian, $ \lambda_1 \geq  \lambda_2 \geq \lambda_3$ are calculated. The $\lambda_3$ field is shown in the top right panel of \autoref{fig:labelsfilter}.

@@

\begin{figure}
\begin{minipage}[t]{0.99\linewidth}
 \centering\includegraphics[height=8.5cm]{Chapter5/Source_v2/fig3.pdf} 
\end{minipage}\hfill
\caption{ Detection of potential halo candidates in the multistream field: algorithm of segmentation and filtering are illustrated in a smaller slice of $40 h^{-1} Mpc \times 40 h^{-1} Mpc$ slice of the simulation box. Top left figure shows the multistream field of the slice. Voids (white) are the regions with $n_{str} =1$, rest are non-void structures. Blue patches within the structure (gray) are the regions with gravitational collapses in more than one direction, i.e., $n_{str} \geq 7$. Top Right figure shows the smallest eigenvalue $\lambda_3$ field. The value of $\lambda_3$ is close to 0 in most of the regions (yellow), including the voids. Regions with  $\lambda_3 > 0$ and $n_{str} > 1$, are isolated (black spots) using image segmentation techniques. Bottom left panel shows the filtering scheme: the red patches do not have maxima of $n_{str} \geq 7$ in the regions, hence are filtered out. The remaining potential halo regions with more than 20 particles are shown in the bottom right panel.}
\label{fig:labelsfilter}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{Chapter5/Source_v2/fig4.pdf} 
\end{minipage}\hfill
\caption{PDF of highest $\lambda_1$, $\lambda_2$ and $\lambda_3$ values in each of 4492 haloes detected by out algorithm. The peaks of the PDF are in the range 1-10. Shaded regions represent 1$\sigma$ error. }
\label{fig:maxL3}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{Chapter5/Source_v2/fig5.pdf} 
\end{minipage}\hfill
\caption{The distribution of $\lambda_3$ in the simulation box of $128^3$ particles and multistream field of refinement factor $l_l/l_d = 2$. Three smoothing scales are shown. }
\label{fig:Eval3Smooth}
\end{figure}

@@

\begin{table}
  \caption{Number of $\lambda_3$-haloes identified at smoothing of $n_{str}(\mathbf{x})$ at different scales.}
\begin{tabular}{|l|r|r|r|}
\hline
$N_p$  &  $0.20h^{-1} \text{Mpc}$ & $0.39 h^{-1} \text{Mpc}$ & $0.78 h^{-1} \text{Mpc}$ \\ \hline
$128^3$   & $5181$  &  $4492$ & $2923$ \\ \hline
$256^3$   & $27929$  & $18221$ & $7897$ \\ \hline

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{Chapter5/Source_v2/fig6.pdf} 
\end{minipage}\hfill
\caption{Top panel shows Halo Mass functions of the potential $\lambda_3$-haloes in the multistream field refinement factor $l_l/l_d = 2$ with various smoothing scales. Simulation box has $256^3$ particles. Lower panel shows the deviation of the each halo mass function with respect to their average.}
\label{fig:hmfSmooth}
\end{figure}

@@

Environmental dependence of the haloes results in various multistream values for the halo core. Theoretical toy models of halo formation, such as the tetrahedral collapse model \cite{Neyrinck2016} describes a three-dimensional halo with four filaments accreting mass into it, has 15 stream crossings.  
\cite{Ramachandra2015} have previously showed that a high threshold of $n_{str} \geq 90$ is equivalent of virial density of $\rho_{vir} = 200$, and filters most of the large haloes above $10^{13} M_{\odot}$.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{Chapter5/Source_v2/fig7.pdf} 
\end{minipage}\hfill
\caption{Halo Mass functions of the potential $\lambda_3$-haloes in the multistream field refinement factor $l_l/l_d = 2$ (in Simulation box with $N_p = 256^3$) with various thresholds on local maxima of $n_{str}$ within the halo. Lower panel shows the deviation of the each halo mass function with respect to their average.}
\label{fig:hmfEth}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{Chapter5/Source_v2/fig8.pdf} 
\end{minipage}\hfill
\caption{Maximum, minimum and median of $n_{str}$ in each of 4492 halo candidates. The closed contours of haloes encompass a wide range of multistream values. None of the haloes are in contact with the void region, since lowest value of min$(n_{str})$ is 3. Shaded regions are the 1$\sigma$ absolute errors in the number of $\lambda_3$-haloes. }
\label{fig:minmaxstr}
\end{figure}

@@

However, the $\lambda_3$-halo boundary is different from any constant multistream contour. That is, from the function $n_{str}(\mathbf{x})$, convex regions in the four-dimensional function space $(-n_{str}, x, y, z)$ are projected onto three-dimensional co-ordinate space using eigenvalues. This is different from projecting `iso-multistream' slice onto three-dimensional co-ordinate space. Appendix \ref{appendix:Eigen} illustrates the difference in the two approaches for a one-dimensional function.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=7.cm]{Chapter5/Source_v2/fig9.pdf} 
\end{minipage}\hfill
\caption{Multistream environment of a $\lambda_3$-halo. The contours represent regions with 3 different multistream values: Outermost $n_{str} \geq 25$ (gray) is tubular, The blue region has $n_{str} \geq 75$. The inner region is highly non-linear with $n_{str} \geq 200$.  The black dots represent the mass particles belonging to a $\lambda_3$-halo, as detected by our algorithm.}
\label{fig:l3Nst}
\end{figure}

The particles in a massive $\lambda_3$-halo shown in \autoref{fig:l3Nst} form a spheroidal structure. The number of particles in similar massive haloes are in the order of $10^3 - 10^4$ particles. For instance, the most massive halo in the simulation (with $N_p = 128^3$) has $5593$ particles. We have chosen a minimum threshold of 20 particles, which is an artificial parameter (may be cooked up or ad hoc) used by most halo finder methods. Majority of the $\lambda_3$-haloes have low number of particles; median of number of particles per halo is $49$. Each particle in this simulation is approximately $ 3.65 \times 10^{10} h^{-1} M_{\odot}$. Hence the halo mass range varies in the order of $10^{11} M_{\odot}$ to $10^{14} M_{\odot}$. Combined mass of all the $\lambda_3$-halo candidates is about $31$ per cent of the total mass in the simulation. In contrast, the haloes occupy just $0.3$ per cent of the total volume. Thus the $\lambda_3$-haloes are extremely dense structures. Further analysis of halo mass function of $\lambda_3$-haloes and comparison with AHF- and FOF-haloes is done in the Section \ref{sub:compareHalo}.

@@

The halo mass functions from all three finders are shown in \autoref{fig:hmf}. For smaller haloes of order of $10^{13} M_{\odot}$, our method predicts a slightly higher number of haloes than FOF and AHF. For the most massive haloes of mass around $10^{14} M_{\odot}$, number of $\lambda_3$-haloes is fewer than the other 2 methods, albeit around the range of error of AHF-haloes.

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{Chapter5/Source_v2/fig10.pdf} 
\end{minipage}\hfill
\caption{Halo Mass functions from AHF-, FOF- and $\lambda_3$-haloes. The AHF-haloes are fewer than FOF- or $\lambda_3$-haloes. The number of haloes above a mass threshold are binned and taken along vertical axis, normalized to simulation box volume. Error of 1$\sigma$ is shown in shaded region. }
\label{fig:hmf}
\end{figure}

By observing some of the massive haloes, like the one in \autoref{fig:FinderCompareAll}, we find that the $\lambda_3$-halo particles are within AHF- or FOF-halo region. This is generally observed in other massive haloes too: the large $\lambda_3$-haloes have fewer particles than corresponding AHF- or FOF-haloes. For haloes greater than $10^{14} M_{\odot}$, $\lambda_3$-haloes have boundaries slightly within the AHF virial radius. Without unbinding, the FOF-haloes can be very large compared to other methods, as seen in \autoref{fig:FinderCompareAll}. This results in a deviation in the $\lambda_3$-halo mass function (\autoref{fig:hmf}) from the other two methods over halo mass of $10^{14} M_{\odot}$. Further discussion of size of the detected $\lambda_3$-haloes in the context of smoothing of the multistream is done in Section \ref{sub:Smooth}.

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering
 \includegraphics[width=8.cm]{Chapter5/Source_v2/fig11.pdf}
\end{minipage}\hfill
\caption{A large halo that was detected by AHF (top, red), our geometric analysis in the multistream field (centre, green) and FOF (bottom, blue). Halo boundary differs for each halo finder method. AHF detects particles within a sphere of virial radius. FOF-halo is irregularly shaped. $\lambda_3$-halo particles are in a non-spherical, yet compact structure. }
\label{fig:FinderCompareAll}
\end{figure}

@@

\includegraphics[width=7.7cm]{Chapter5/Source_v2/fig12a.pdf}  
\includegraphics[width=7.7cm]{Chapter5/Source_v2/fig12b.pdf}  
\includegraphics[width=7.7cm]{Chapter5/Source_v2/fig12c.pdf}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering
 \includegraphics[width=8.cm]{Chapter5/Source_v2/fig13.pdf}
\end{minipage}\hfill
\caption{Mass fraction of haloes $f_h$ (in per cent) as a detected by the three finders. Each circle represents fractions of mass of all halo particles (AHF, FOF or $\lambda_3$) in the total mass of the simulation box. The particles concurrently detected as belonging to haloes by different frameworks are shown in the intersections. }
\label{fig:HaloFinderMF}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter5/Source_v2/fig14.pdf} 
\end{minipage}\hfill
\caption{ Percentage of haloes detected (in the simulation with $N_p = 256^3$) that are embedded in the non-excursion set (top panel), largest excursion set segment (center panel) and the rest of excursion set regions (bottom panel). Multistream haloes, AHF-haloes and FOF-haloes are shown in green, red and blue respectively. Vertical dashed line at $n_{str} = 17$ is where percolation transition occurs. }
\label{fig:HaloFilAll}
\end{figure}

@@

The multistream field can have only integral  values, moreover these values must be odd numbers practically at every point, because the number of streams can be even only on a set of measure zero. It means that in numerical simulations  even values may occur on extremely rare occasions. 
We have analysed functional variation of the scalar field $-n_{str}(\mathbf{x})$ using Hessian eigenvalues. The Hessian analysis is generally done for inherently continuous fields since it requires  evaluation of the second derivatives. % of  $n_{str}(\mathbf{x})$. }
Geometries of web structures unveiled by Hessian signatures of smoothed density fields (such as \citealt{Sousbie2008a}, \citealt{Aragon-Calvo2007}, \citealt{Aragon-Calvo2010a}, \citealt{Cautun2014a}, \citealt{Bond2010a} and many more), tidal shear or velocity shear tensor (\citealt{Hahn2007}, \citealt{Forero-Romero2009a}, \citealt{Hoffman2012a}, \citealt{Hoffman2012a}, \citealt{Libeskind2013}, \citealt{Cautun2014a} etc.) and observational data from surveys (\citealt{Sousbie2008a}, \citealt{Bond2010a}, \citealt{Bond2010b}, \citealt{Pahwa2016} etc.).

@@

\begin{comment}
\bibliographystyle{mnras}
\bibliography{library}

>>>>

\begingroup
\let\clearpage\relax
%\tableofcontents
\endgroup
\newpage
%-------------------------------------------------------------------------------------------------------------------------------

@@

\begin{figure*} 
\centering\includegraphics[height=7cm]{fig1.pdf} 
\caption{Dynamical collapse of dark matter in one-dimensional universe: Top panels show the $(\bmath{p}, \bmath{x})$ phase-space manifold of the dark matter sheet at redshifts $z_1$, $z_2$, $z_3$ and $z = 0$. Dots represent the dark matter particles. The momentum values are chosen at arbitrary scales. Bottom panels show the corresponding multistream field multistream field $n_{str}(\bmath{x}, z)$ (red) and density field $\rho(\bmath{x}, z)$ (gray). }
\label{fig:1d}
\end{figure*}

We begin with a simple illustration showing the formation of a few haloes in a one-dimensional simulation. Dark matter clustering in a (1+1)-dimensional phase-space $(\bmath{p}, \bmath{x})$ (where $p$ is the momentum and $x$ is the co-moving Eulerian coordinate) at four successive time steps is shown in the top panels of \autoref{fig:1d}. The lower panels show the corresponding multistream field (\citealt{Shandarin2012} and \citealt{Abel2012}) $n_{str}(\bmath{x}, z)$ (red) and density field $\rho(\bmath{x}, z)$ (gray). At $z_1$ (left-most panel), velocity is single-valued in Eulerian co-ordinates shown, except at a small three-stream region near $\bmath{x} = 5\pi/4$. This is the first instance of multistreaming in the region, which was previously had $n_{str} = 1$ throughout. The interface of $n_{str} = 1$ and $n_{str} = 3$ regions is also the location of the first caustic. On the other hand, the density calculated at a high resolution shows variations, even in the mono-streaming regions. The variations are especially more pronounced around  the caustic (near $\bmath{x} = 5\pi/4$).

@@

\begin{figure} 
\centering\includegraphics[width=8cm]{fig2.pdf} 
\caption{Multistream field contours: The multistream field is calculated at 8 times the native resolution. Each inner convex blobs (red) surround local multistream maxima inside. Surrounding outer shell(blue) is not convex throughout the surface, and the outermost gray multistream surface displays a filamental geometry.}
\label{fig:full}
\end{figure}

@@

The emphasis of this paper is to demonstrate the use of multistream field in identifying potential dark matter haloes, and not a full statistical analysis of halo properties. For this purpose, we have run simulations at two different mass resolutions (number of particles $N_p = 128^3$ and $256^3$, and respective mass of particles, $m_p = 3.65 \times 10^{10} h^{-1} M_{\sun}$ and $ 4.57 \times 10^{9} h^{-1} M_{\sun}$), with the same periodic side length $L = 100 h^{-1} Mpc$. The gravitational softening length $\epsilon = 20 h^{-1} kpc$ and $10 h^{-1} kpc$ for low and high resolution simulation respectively. The initial conditions are generated by {MUSIC} \citep{Hahn2011a} with the transfer function from \cite{Eisenstein1998a} at a redshift of $z_{ini}= 80$. The $\Lambda$CDM cosmological simulation is run using {GADGET-2} (\citealt{Springel2005} and \citealt{Springel2001}) is similar to the ones used in \cite{Ramachandra2017}. The cosmological parameters used in the simulation are $\Omega_{m}= 0.276$, $\Omega_{\Lambda}= 0.724$, the Hubble parameter, $h = 0.703$, the power spectrum normalization, $\sigma_8 = 0.811$ and the spectral index $n_s= 0.961$.

@@

We intend to identify haloes in the $n_{str}(\bmath{x})$ field instead of using just the position coordinate data. While the eigenvalue analysis itself is done at a chosen time, the multistream field inherently has data from six-dimensional Lagrangian space $(\bmath{q}, \bmath{x})$ that contains the full dynamical information,similar to the phase-space sheet albeit in a different form. Dynamical history that is embedded in the multistream field is crucial in understanding the strength of gravitational binding of the particles. 
A physically motivated distinction between void and gravitationally collapsed regions -- voids are the regions with a single stream -- is a unique feature of multistream analysis (\citealt{Shandarin2012} and \citealt{Ramachandra2017}). Thus the haloes detected from local maxima of the $n_{str}$ field can be ensured to be away from the mono-streaming voids. Methods based on linking-length or density fields may not be able to ensure that  all the particles in haloes are away from voids (as shown for FOF haloes in \citealt{Ramachandra2017}) .

@@

Hessian matrix $\mathbfss{H}(f)$ of a scalar field $f$ involves local second-order variations in three orthogonal directions. Each element of the Hessian matrix $\mathbfss{H}_{ij}(f)$ (where $i$ and $j$ can be any of $x$, $y$ or $z$ directions) is given by \autoref{eq:Hess}.

\begin{equation}
\label{eq:Hess}
 \mathbfss{H}_{ij}(f) = \frac{\partial^2 f}{\partial x_i \partial x_j}
\end{equation}

@@

(i) Every element of Hessian matrices $\mathbfss{H}(-n_{str})$, and consequently the eigenvalues $\lambda_i$'s are zero in single-stream voids. Even if the multistream field is a smoothed, the eigenvalues peak at zero. This property is unique to multistream fields. Eigenvalues of Hessians of density \cite{Aragon-Calvo2007}, velocity shear tensor \cite{Libeskind2013} do not peak at zero, and the eigenvalues of deformation tensor are negative in voids as a result of continuity equation (shown in Zel’dovich formalism as well).

@@

{Our goal is to isolate the locations of convex geometries in the multistream flow field. Prospective regions of the halo locations in the web structure are selected by positive definite condition of the Hessian $\mathbfss{H}(-n_{str})$: $\lambda_1 > 0$, $\lambda_2 > 0$ and  $\lambda_3 > 0$, or simply the smallest of the eigenvalues, $\lambda_3 > 0$. We also filter out the regions if the multistream values inside do not suggest gravitational collapse into haloes. The sequence of our halo detection framework is listed below:

@@

\item Second order variations of the smoothed $-n_{str}(\bmath{x})$ is computed at each point in the field. This gives symmetric Hessian matrices for this field whose eigenvalues are real. Ordered eigenvalues of the Hessian, $ \lambda_1 \geq  \lambda_2 \geq \lambda_3$ are calculated. The $\lambda_3$ field is shown in the top right panel of \autoref{fig:labelsfilter}.

@@

\begin{figure}
\begin{minipage}[t]{0.99\linewidth}
 \centering\includegraphics[height=8.5cm]{fig3.pdf} 
\end{minipage}\hfill
\caption{ Detection of potential halo candidates in the multistream field: algorithm of segmentation and filtering are illustrated in a smaller slice of $40 h^{-1} Mpc \times 40 h^{-1} Mpc$ slice of the simulation box. Top left figure shows the multistream field of the slice. Voids (white) are the regions with $n_{str} =1$, rest are non-void structures. Blue patches within the structure (gray) are the regions with gravitational collapses in more than one direction, i.e., $n_{str} \geq 7$. Top Right figure shows the smallest eigenvalue $\lambda_3$ field. The value of $\lambda_3$ is close to 0 in most of the regions (yellow), including the voids. Regions with  $\lambda_3 > 0$ and $n_{str} > 1$, are isolated (black spots) using image segmentation techniques. Bottom left panel shows the filtering scheme: the red patches do not have maxima of $n_{str} \geq 7$ in the regions, hence are filtered out. The remaining potential halo regions with more than 20 particles are shown in the bottom right panel.}
\label{fig:labelsfilter}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{fig4.pdf} 
\end{minipage}\hfill
\caption{PDF of highest $\lambda_1$, $\lambda_2$ and $\lambda_3$ values in each of 4492 haloes detected by out algorithm. The peaks of the PDF are in the range 1-10. Shaded regions represent 1$\sigma$ error. }
\label{fig:maxL3}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{fig5.pdf} 
\end{minipage}\hfill
\caption{The distribution of $\lambda_3$ in the simulation box of $128^3$ particles and multistream field of refinement factor $l_l/l_d = 2$. Three smoothing scales are shown. }
\label{fig:Eval3Smooth}
\end{figure}

@@

\begin{table}
  \caption{Number of $\lambda_3$-haloes identified at smoothing of $n_{str}(\bmath{x})$ at different scales.}
\begin{tabular}{|l|r|r|r|}
\hline
$N_p$  &  $0.20h^{-1} \text{Mpc}$ & $0.39 h^{-1} \text{Mpc}$ & $0.78 h^{-1} \text{Mpc}$ \\ \hline
$128^3$   & $5181$  &  $4492$ & $2923$ \\ \hline
$256^3$   & $27929$  & $18221$ & $7897$ \\ \hline

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{fig6.pdf} 
\end{minipage}\hfill
\caption{Top panel shows Halo Mass functions of the potential $\lambda_3$-haloes in the multistream field refinement factor $l_l/l_d = 2$ with various smoothing scales. Simulation box has $256^3$ particles. Lower panel shows the deviation of the each halo mass function with respect to their average.}
\label{fig:hmfSmooth}
\end{figure}

@@

Environmental dependence of the haloes results in various multistream values for the halo core. Theoretical toy models of halo formation, such as the tetrahedral collapse model \cite{Neyrinck2016} describes a three-dimensional halo with four filaments accreting mass into it, has 15 stream crossings.  
\cite{Ramachandra2015} have previously showed that a high threshold of $n_{str} \geq 90$ is equivalent of virial density of $\rho_{vir} = 200$, and filters most of the large haloes above $10^{13} M_{\sun}$.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{fig7.pdf} 
\end{minipage}\hfill
\caption{Halo Mass functions of the potential $\lambda_3$-haloes in the multistream field refinement factor $l_l/l_d = 2$ (in Simulation box with $N_p = 256^3$) with various thresholds on local maxima of $n_{str}$ within the halo. Lower panel shows the deviation of the each halo mass function with respect to their average.}
\label{fig:hmfEth}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{fig8.pdf} 
\end{minipage}\hfill
\caption{Maximum, minimum and median of $n_{str}$ in each of 4492 halo candidates. The closed contours of haloes encompass a wide range of multistream values. None of the haloes are in contact with the void region, since lowest value of min$(n_{str})$ is 3. Shaded regions are the 1$\sigma$ absolute errors in the number of $\lambda_3$-haloes. }
\label{fig:minmaxstr}
\end{figure}

@@

However, the $\lambda_3$-halo boundary is different from any constant multistream contour. That is, from the function $n_{str}(\bmath{x})$, convex regions in the four-dimensional function space $(-n_{str}, x, y, z)$ are projected onto three-dimensional co-ordinate space using eigenvalues. This is different from projecting `iso-multistream' slice onto three-dimensional co-ordinate space. Appendix \ref{appendix:Eigen} illustrates the difference in the two approaches for a one-dimensional function.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=7.cm]{fig9.pdf} 
\end{minipage}\hfill
\caption{Multistream environment of a $\lambda_3$-halo. The contours represent regions with 3 different multistream values: Outermost $n_{str} \geq 25$ (gray) is tubular, The blue region has $n_{str} \geq 75$. The inner region is highly non-linear with $n_{str} \geq 200$.  The black dots represent the mass particles belonging to a $\lambda_3$-halo, as detected by our algorithm.}
\label{fig:l3Nst}
\end{figure}

The particles in a massive $\lambda_3$-halo shown in \autoref{fig:l3Nst} form a spheroidal structure. The number of particles in similar massive haloes are in the order of $10^3 - 10^4$ particles. For instance, the most massive halo in the simulation (with $N_p = 128^3$) has $5593$ particles. We have chosen a minimum threshold of 20 particles, which is an artificial parameter (may be cooked up or ad hoc) used by most halo finder methods. Majority of the $\lambda_3$-haloes have low number of particles; median of number of particles per halo is $49$. Each particle in this simulation is approximately $ 3.65 \times 10^{10} h^{-1} M_{\sun}$. Hence the halo mass range varies in the order of $10^{11} M_{\sun}$ to $10^{14} M_{\sun}$. Combined mass of all the $\lambda_3$-halo candidates is about $31$ per cent of the total mass in the simulation. In contrast, the haloes occupy just $0.3$ per cent of the total volume. Thus the $\lambda_3$-haloes are extremely dense structures. Further analysis of halo mass function of $\lambda_3$-haloes and comparison with AHF- and FOF-haloes is done in the Section \ref{sub:compareHalo}.

@@

The halo mass functions from all three finders are shown in \autoref{fig:hmf}. For smaller haloes of order of $10^{13} M_{\sun}$, our method predicts a slightly higher number of haloes than FOF and AHF. For the most massive haloes of mass around $10^{14} M_{\sun}$, number of $\lambda_3$-haloes is fewer than the other 2 methods, albeit around the range of error of AHF-haloes.

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=8.cm]{fig10.pdf} 
\end{minipage}\hfill
\caption{Halo Mass functions from AHF-, FOF- and $\lambda_3$-haloes. The AHF-haloes are fewer than FOF- or $\lambda_3$-haloes. The number of haloes above a mass threshold are binned and taken along vertical axis, normalized to simulation box volume. Error of 1$\sigma$ is shown in shaded region. }
\label{fig:hmf}
\end{figure}

By observing some of the massive haloes, like the one in \autoref{fig:FinderCompareAll}, we find that the $\lambda_3$-halo particles are within AHF- or FOF-halo region. This is generally observed in other massive haloes too: the large $\lambda_3$-haloes have fewer particles than corresponding AHF- or FOF-haloes. For haloes greater than $10^{14} M_{\sun}$, $\lambda_3$-haloes have boundaries slightly within the AHF virial radius. Without unbinding, the FOF-haloes can be very large compared to other methods, as seen in \autoref{fig:FinderCompareAll}. This results in a deviation in the $\lambda_3$-halo mass function (\autoref{fig:hmf}) from the other two methods over halo mass of $10^{14} M_{\sun}$. Further discussion of size of the detected $\lambda_3$-haloes in the context of smoothing of the multistream is done in Section \ref{sub:Smooth}.

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering
 \includegraphics[width=8.cm]{fig11.pdf}
\end{minipage}\hfill
\caption{A large halo that was detected by AHF (top, red), our geometric analysis in the multistream field (centre, green) and FOF (bottom, blue). Halo boundary differs for each halo finder method. AHF detects particles within a sphere of virial radius. FOF-halo is irregularly shaped. $\lambda_3$-halo particles are in a non-spherical, yet compact structure. }
\label{fig:FinderCompareAll}
\end{figure}

@@

\includegraphics[width=7.7cm]{fig12a.pdf}  
\includegraphics[width=7.7cm]{fig12b.pdf}  
\includegraphics[width=7.7cm]{fig12c.pdf}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering
 \includegraphics[width=8.cm]{fig13.pdf}
\end{minipage}\hfill
\caption{Mass fraction of haloes $f_h$ (in per cent) as a detected by the three finders. Each circle represents fractions of mass of all halo particles (AHF, FOF or $\lambda_3$) in the total mass of the simulation box. The particles concurrently detected as belonging to haloes by different frameworks are shown in the intersections. }
\label{fig:HaloFinderMF}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig14.pdf} 
\end{minipage}\hfill
\caption{ Percentage of haloes detected (in the simulation with $N_p = 256^3$) that are embedded in the non-excursion set (top panel), largest excursion set segment (center panel) and the rest of excursion set regions (bottom panel). Multistream haloes, AHF-haloes and FOF-haloes are shown in green, red and blue respectively. Vertical dashed line at $n_{str} = 17$ is where percolation transition occurs. }
\label{fig:HaloFilAll}
\end{figure}

@@

The multistream field can have only integral  values, moreover these values must be odd numbers practically at every point, because the number of streams can be even only on a set of measure zero. It means that in numerical simulations  even values may occur on extremely rare occasions. 
We have analysed functional variation of the scalar field $-n_{str}(\bmath{x})$ using Hessian eigenvalues. The Hessian analysis is generally done for inherently continuous fields since it requires  evaluation of the second derivatives. % of  $n_{str}(\bmath{x})$. }
Geometries of web structures unveiled by Hessian signatures of smoothed density fields (such as \citealt{Sousbie2008a}, \citealt{Aragon-Calvo2007}, \citealt{Aragon-Calvo2010a}, \citealt{Cautun2014a}, \citealt{Bond2010a} and many more), tidal shear or velocity shear tensor (\citealt{Hahn2007}, \citealt{Forero-Romero2009a}, \citealt{Hoffman2012a}, \citealt{Hoffman2012a}, \citealt{Libeskind2013}, \citealt{Cautun2014a} etc.) and observational data from surveys (\citealt{Sousbie2008a}, \citealt{Bond2010a}, \citealt{Bond2010b}, \citealt{Pahwa2016} etc.).

@@

\bibliographystyle{mnras}
\bibliography{library}
